---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(sc,"SELECT COUNT(DISTINCT product_id) AS products
                FROM order_products__prior_tbl")
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity

```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```


```{r}
products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```


```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?

```{r}
ex1 <- "select order_dow, count(*) as n 
        from orders_tbl 
        group by order_dow
        order by count(*) desc
        LIMIT 1"
dbGetQuery(sc, ex1) #0 -> diumenge
```

* The most popular hour of doing order? 

```{r}
ex2 <- "select order_hour_of_day, count(*) as n 
        from orders_tbl 
        group by order_hour_of_day
        order by count(*) desc
        LIMIT 1"
dbGetQuery(sc, ex2)
```

* The most reordered products?

```{r}
ex3 <- "select product_id, sum(reordered)
        from order_products__prior_tbl
        group by product_id
        order by sum(reordered) desc
        LIMIT 5"
dbGetQuery(sc, ex3)
```

* The less reorderd product?

```{r}
ex4 <- "select product_id, sum(reordered)
        from order_products__prior_tbl
        group by product_id
        order by sum(reordered)
        LIMIT 10"
dbGetQuery(sc, ex4)
```

* Plot histogram of number of products in an order.

```{r}
ex5 <- "select order_id, count(*)
        from order_products__prior_tbl
        group by order_id"
data <- dbGetQuery(sc, ex5) 
```

* What is the most popular department? 

## Whisker, magrittr

Whisker ens permet posar el valor de variables dins un string.
Si var=list(nom="Marta"), whisker.render("Hola {{nom}}",var) retorna "Hola Marta"

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat #veiem que agafa el valor 24852
```


```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).
```{r}
dbGetQuery(sc, 
"SELECT *
FROM products_tbl
limit 20") 
```

* For given day, rank department popularity.

## More on orders (Window Functions)

Per exemple si tenim aquesta taula i volem saber per al total d'estudiants de cada departament, quants corresponen a cada any -> **window function**

```{r}
data <- data.frame(department=c(1,1,1,2,2), year=c(2010,2011,2012,2010,2011), n_student=c(20,21,25,13,18))
data_to_spark <- copy_to(sc, data, "data_s", overwrite = TRUE) #passem dades a spark

dbGetQuery(sc, 
"SELECT department, n_student/sum(n_student) OVER(partition by department) as 
FROM data_s") 
```

```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```

Per a casa usuari (agafem només usuaris fins al 10), agafem la primera ordre que va fer i la guardem a la columna first_order.

```{r}
dbGetQuery(sc, 
"SELECT user_id, order_id, FIRST_VALUE(order_id) OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```

Perquè surti només una línia per usuari:

```{r}
dbGetQuery(sc, 
"SELECT distinct user_id, FIRST_VALUE(order_id) OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
ORDER BY user_id
")
```

Ara volem buscar el producte més comú en les cistelles grans:

* Product that is usually orders in numerous baskets

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* Could modify the query and add prduct name?

```{r}
"
SELECT pn.product_id, product_name
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn LEFT JOIN PRODUCTS_TBL t2 on pn.product_id=t2.product_id
GROUP BY pn.product_id, product_name
ORDER BY avg_products_number DESC
LIMIT 20
" %>%
dbGetQuery(sc, .)
```

* Product that is usually orders in small baskets

```{r}
"
SELECT pn.product_id, product_name
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn LEFT JOIN PRODUCTS_TBL t2 on pn.product_id=t2.product_id
GROUP BY pn.product_id, product_name
ORDER BY avg_products_number
LIMIT 20
" %>%
dbGetQuery(sc, .)
```

* For 5 first users and all their orders find days since first order.

```{r}
"
SELECT order_id,user_id,order_number,days_since_prior_order, sum(days_since_prior_order) over (partition by user_id order by order_number, user_id rows unbounded preceding) as dies_dsd_1
FROM orders_tbl
WHERE user_id<=5
ORDER BY user_id, order_number
" %>%
dbGetQuery(sc, .)
```

